---
title: "Basic Stat Lab"
format: html
editor: visual
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3.5, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
library(stats)
```

# Basic Statistics Classwork/Lab

1)  In Hong Kong, human male height is approximately normally distributed with mean 171.5 cm and standard deviation 5.5 cm. (Use and show R code to produce answers for a -- e) You can use any method for part f.\
    a) What proportion of the Hong Kong population is between 170 cm and 180 cm?

    ```{r}
    #if X∼N(μ,σ2) and Z=(X−μ)/σ, then Z∼N(0,1).
    #The distribution of X is Normal(171.5(5.5)^2) then the Probability of the interval is equal to the difference between P(X ≤ 170) and P(X ≤ 160)
    print(pnorm(180,171.5,5.5) - pnorm(170,171.5,5.5))
    ```

    b\) What proportion of the Hong Kong population is less that 165 cm ?

    ```{r}
    #P(X ≤ 165)
    print(1 - pnorm(165,171.5,5.5))
    ```

    c\) What proportion of the Hong Kong population is greater than 175 ?

    ```{r}
    #P(X ≤ 175)
    print(1 - pnorm(175,171.5,5.5))
    ```

    d\) The proportion .25 is less than what male height value ?

    ```{r}
    #P(X ≤ x) = .25
    #.25 -> z-score ~ -.67 = (X - 171.5)/5.5
    print(qnorm(.25, 171.5, 5.5))
    ```

    e\) The proportion .85 is greater than what male height value ?

    ```{r}
    #P(X ≤ x) = .85
    #.85 -> z-score ~ 1.04 = (X - 171.5)/5.5
    print(qnorm(.85, 171.5, 5.5))
    ```

    f\) The proportion of .68 is between what two male height values ?

    ```{r}
    #In a normally distributed sample 68% of all the data is 1 standard deviation between the mean.
    print(171.5 + 5.5)
    print(171.5 - 5.5)
    print("68% of men fall between the height of 166 cm and 177 cm")
    ```

2)  Use and show R code to determine the height of the standard normal curve at a Z value of 2.5.

    ```{r}
    #2.5 = (X - 171.5)/5.5
    #X = (2.5*(5.5) + 171.5)
    dnorm((2.5*(5.5) + 171.5), 171.5, 5.5)
    ```

3)  Use and show R code to determine the height of the standard normal curve at a Z value of .4.

    ```{r}
    #.4 = (X - 171.5)/5.5
    #X = (.4*(5.5) + 171.5)
    dnorm((.4*(5.5) + 171.5), 171.5, 5.5)
    ```

4)  Use and show R code to determine the height of a t distribution curve for a t value of 1.5 with 3 degrees of freedom.

    ```{r}
    dt(x = 1.5, df = 3)
    ```

5)  Use and show R code to determine the height of a t distribution curve for a t value of -2 with 2 degrees of freedom.

    ```{r}
    dt(x = -2, df = 2)
    ```

6)  For a t distribution with 3 degrees of freedom, use and show R code that will find the proportion less than 2.5.

    ```{r}
    print(pt(2.5, df = 3))
    ```

7)  For a t distribution with 1 degree of freedom, use and show R code that will find the proportion that is greater than 1.75.

    ```{r}
    print(pt(1.75, 1))
    ```

8)  For a t distribution with 2 degrees of freedom, use and show R code that will find the value immediately above a proportion of .355.

    ```{r}
    print(qt(.355, df = 2, lower.tail = FALSE))
    ```

9)  An educator believes that new directed reading activities in the classroom will help elementary school pupils improve some aspects of their ability. She arranges for a third -grade class of 23 students to take part in these activities for an eight-week period. A control classroom of 23 third graders follows the same curriculum without the activities. At the end of the eight weeks, all students are given a Degree of Reading Power (DRP) test, which measures the aspects of reading ability that the treatment is designed to improve. The sample data performance results are provided below;

-   Treatment Group: 24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70

-   Control Group: 42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42

-   Design and execute a two sample t test.\
     a. State the appropriate null and alternative hypotheses.

    **H~0~:** µ~1~ = µ~2~ (the two population means are equal)

    **H~A~:** µ~1~ ≠µ~2~ (the two population means are *not* equal)

    b)  Use and show R code to produce the p value and the confidence interval

        ```{r}
        test_group <- c(24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70)
        control_group <- c(42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42)

        t.test(test_group, control_group)
        print("")
        ```

    c)  Determine if you should reject or fail to reject the null hypothesis using the p value and the confidence interval.

        Because the p-value of our test **(.01054)** is less than 0.05, we reject the null hypothesis.
