---
title: "3_23 Classwork"
author: "Connor Rempe"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(Sleuth3)
```

1.In Hong Kong, human male height is approximately normally distributed with mean 171.5 cm and standard deviation 5.5 cm. (Use and show R code to produce answers for a – e) You can use any method for part f.

```{r}
l <- 171.5 - 5*5.5
u <- 171.5 + 5*5.5
x <- seq(l,u, length = 100)
length(x)
y <- dnorm(x = x, mean = 171.5, sd = 5.5)
```

a) What proportion of the Hong Kong population is between 170 cm and 180 cm?
```{r}
pnorm(180, mean = 171.5, sd = 5.5) - pnorm(170, mean = 171.5, sd = 5.5)
```
54 percent of men in Hong Kong are between 170cm and 180cm

b) What proportion of the Hong Kong population is less that 165 cm ?
```{r}
pnorm(165, mean = 171.5, sd = 5.5) 
```
11 percent of men in Hong Kong are less than 165cm tall. 

```{r}
df_1 <- data.frame(x=c(min(x),x[x < 165],165), 
                  y = c(0, y[x < 165], 0))
qplot(x,y,geom = "line", ylab = "f(x)")+
  geom_polygon(data = df_1, mapping = aes(x = x, y = y), fill= "blue", alpha = 1/4)
```


c) What proportion of the Hong Kong population is greater than 175 ?
```{r}
pnorm(175, mean = 171.5, sd = 5.5, lower.tail = FALSE)
```
26 percent of the men in Hong Kong are taller than 175cm. 
```{r}
df_1 <- data.frame(x=c(175,x[x>175],max(x)), 
                  y = c(0, y[x >175], 0))
qplot(x,y,geom = "line", ylab = "f(x)")+
  geom_polygon(data = df_1, mapping = aes(x = x, y = y), fill= "blue", alpha = 1/4)
```


d) The proportion .25 is less than what male height value ?
```{r}
x_25 <- qnorm(.25, mean = 171.5, sd= 5.5)
x_25
```
0.25 percent of men in Hong Kong are shorter than 167.8 cm. 
```{r}
df_1 <- data.frame(x = c(min(x),x[x < x_25],x_25),
                  y = c(0, y[x < x_25], 0))
qplot(x,y,geom = "line", ylab = "f(x)")+
  geom_polygon(data = df_1, mapping = aes(x = x, y = y), 
               fill= "blue", alpha = 1/4)+
  annotate(geom = "text",x = 167, y = 0.01, label = 0.25, color = "black")
```


e) The proportion .85 is greater than what male height value ?
```{r}
x_85 <- qnorm(.85, mean = 171.5, sd= 5.5, lower.tail = FALSE)
x_85
```

0.85 percent of men in Hong Kong are shorter than 165.8 cm. 
```{r}
df_1 <- data.frame(x = c(x_85,x[x > x_85],max(x)),
                  y = c(0, y[x > x_85], 0))
qplot(x,y,geom = "line", ylab = "f(x)")+
  geom_polygon(data = df_1, mapping = aes(x = x, y = y), 
               fill= "blue", alpha = 1/4)+
  annotate(geom = "text",x = 165, y = 0.01, label = 0.85, color = "black")
```


f) The proportion of .68 is between what two male height values 
```{r}
qnorm(.98, mean = 171.5, sd= 5.5) 
qnorm(.30, mean = 171.5, sd= 5.5)

```
68% of men in Hong Kong are between 168.6 and 182.7 cm tall. 

2.Use and show R code to determine the height of the standard normal curve at a Z value of 2.5.
```{r}
dnorm(x = 2.5, mean = 0, sd = 1)
```
The height of the standard normal curve at a z value of 2.5 is 0.018. 
```{r}
z <- seq(-5,5, length = 100)
Y <- dnorm(x = z, mean = 0, sd = 1)
df_2 <- tibble(z,Y)
```


3.Use and show R code to determine the height of the standard normal curve at a Z value of .4.
```{r}
dnorm(x = 0.4, mean = 0, sd = 1)
```
The height of the standard normal curve at a Z value of 0.4 is 0.368

4.Use and show R code to determine the height of a t distribution curve for a t value of 1.5 with 3 degrees of freedom.
```{r}
t <- seq(-5,5,length= 500)
y_t <- dt(t, df =3)
df_t <- tibble(t,y_t)
```


```{r}
t_1.5<- dt(1.5,3)
t_1.5
```
The height of a t distribution with 3 degrees of freedom at a t value of 1.5 is 0.120

```{r}
qplot(t,y_t, geom = "line", ylab = "f(x)")+
  geom_polygon(data = df_t, mapping = aes(x=t, y= y_t), fill = "white", alpha = 1/4) +
  annotate(geom = "text", x = 1.5, y = t_1.5+0.0015, label = "x", color = "green")+
  annotate(geom = "text", x = 1.5, y = t_1.5+0.0015, label = round(t_1.5,2), color = "red")
```

5.Use and show R code to determine the height of a t distribution curve for a t value of -2 with 2 degrees of freedom.

```{r}
t_1 <- seq(-5,5,length= 500)
y_t1 <- dt(t, df =2)
df_t1 <- tibble(t,y_t)
```

```{r}
t_n2 <- dt(-2, 2)
t_n2
```
The height of a t distribution with 2 degrees of freedom at a t value of -2 is 0.068

```{r}
qplot(t,y_t, geom = "line", ylab = "f(x)")+
  geom_polygon(data = df_t1, mapping = aes(x=t_1, y= y_t1), fill = "white", alpha = 1/4) +
  annotate(geom = "text", x = -2, y = t_n2+0.0015, label = "x", color = "green")+
  annotate(geom = "text", x = -2, y = t_n2+0.0015, label = round(t_n2,2), color = "red")
```



6. For a t distribution with 3 degrees of freedom, use and show R code that will find the proportion less than 2.5.

```{r}
t_2.5<- pt(2.5,3)
t_2.5

```
For a t distribution with 3 degrees of freedom, the proportion less than 2.5 is 96 percent.
```{r}
t <- seq(-6, 6, length = 500)
y_t <-  dt(t, df = 3)
polydf <- tibble(x = c(min(t), t[t <2.5], 2.5, 0), y = c(0,y_t[t <2.5], 0,0))
```

```{r}
qplot(t,y_t, geom = "line", ylab = "f(x)")+
  geom_polygon(data = polydf, mapping = aes(x, y), fill = "blue", alpha = 1/4) 
```


7. For a t distribution with 1 degree of freedom, use and show R code that will find the proportion that is greater than 1.75.
```{r}
1- pt(1.75, 1)
```

For a t distribution with 1 degree of freedom, 16 percent of the distribution is greater than 1.75. 

8. For a t distribution with 2 degrees of freedom, use and show R code that will find the value immediately above a proportion of .355.
```{r}
-qt(0.355, df= 2)
```
For a t distribution with 2 degrees of freedom, the value immediatley above with a proportion of 0.355 is 0.428. 

9. An educator believes that new directed reading activities in the classroom will help elementary school pupils improve some aspects of their ability. She arranges for a third -grade class of 23 students to take part in these activities for an eight-week period. A control classroom of 23 third graders follows the same curriculum without the activities. At the end of the eight weeks, all students are given a Degree of Reading Power (DRP) test, which measures the aspects of reading ability that the treatment is designed to improve. The sample data performance results are provided below;

Treatment Group: 24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70

Control Group: 42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42

Design and execute a two sample t test.
a) State the appropriate null and alternative hypotheses.

H0: There is no relation between new ditected reading activitie in the classroom and improvement of some aspects of elementry school pupils ability
Ha: There is a positive relation between new ditected reading activitie in the classroom and improvement of some aspects of elementry school pupils ability

b)Use and show R code to produce the p value and the confidence interval c) Determine if you should reject or fail to reject the null hypothesis using the p value and the confidence interval.
```{r}
library(broom)
tg <- c(24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70)


cg <- c(42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42)

```

```{r}
t.test(tg,cg, alternative = "greater",mu = 0, var.equal = FALSE)%>%
  tidy()
```
We can reject the null hypothesis because we have a p-value of 0.01054 and 0 is not contained in our confidence interval. 
