---
title: "Basic Stat Lab"
author: "Jay Kang"
format: html
editor: visual
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3.5, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

# Basic Statistics Classwork/Lab

1)  In Hong Kong, human male height is approximately normally distributed with mean 171.5 cm and standard deviation 5.5 cm. (Use and show R code to produce answers for a -- e) You can use any method for part f.\

    a)  What proportion of the Hong Kong population is between 170 cm and 180 cm?\

    ```{r}
    height_low <- ((170 - 171.5) / 5.5)
    height_high <- ((180 - 171.5) / 5.5)
    pnorm(height_high) - pnorm(height_low)
    ```

    b)  What proportion of the Hong Kong population is less that 165 cm ?\

    ```{r}
    pnorm((165 - 171.5) / 5.5)
    ```

    c)  What proportion of the Hong Kong population is greater than 175 ?\

    ```{r}
    1 - pnorm((175 - 171.5) / 5.5)
    ```

    d)  The proportion .25 is less than what male height value ?\

    ```{r}
    qnorm(0.25, 171.5, 5.5)
    ```

    e)  The proportion .85 is greater than what male height value ?\

    ```{r}
    qnorm(0.85, 171.5, 5, 5)
    ```

    f)  The proportion of .68 is between what two male height values ?\

    ```{r}
    height_low <- qnorm(.5 - (0.68/2), 171.5, 5.5)
    height_high <- qnorm(.5 + (0.68/2), 171.5, 5.5)
    print(paste("It is between", height_low, "and", height_high))
    # This is one standard deviation away from the center in both directions using the empirical rule.
    ```

2)  Use and show R code to determine the height of the standard normal curve at a Z value of 2.5.

```{r}
dnorm(2.5)
```

3)  Use and show R code to determine the height of the standard normal curve at a Z value of .4.

```{r}
dnorm(0.4)
```

4)  Use and show R code to determine the height of a t distribution curve for a t value of 1.5 with 3 degrees of freedom.

```{r}
dt(1.5, df = 3)
```

5)  Use and show R code to determine the height of a t distribution curve for a t value of -2 with 2 degrees of freedom.

```{r}
dt(-2, df = 2)
```

6)  For a t distribution with 3 degrees of freedom, use and show R code that will find the proportion less than 2.5.

```{r}
pt(2.5, df = 3)
```

7)  For a t distribution with 1 degree of freedom, use and show R code that will find the proportion that is greater than 1.75.

```{r}
1 - pt(1.75, df = 1)
```

8)  For a t distribution with 2 degrees of freedom, use and show R code that will find the value immediately above a proportion of .355.

```{r}
qt(1 - 0.355, df = 2)
# or
qt(0.355, df = 2, lower.tail = F)
```

9)  An educator believes that new directed reading activities in the classroom will help elementary school pupils improve some aspects of their ability. She arranges for a third -grade class of 23 students to take part in these activities for an eight-week period. A control classroom of 23 third graders follows the same curriculum without the activities. At the end of the eight weeks, all students are given a Degree of Reading Power (DRP) test, which measures the aspects of reading ability that the treatment is designed to improve. The sample data performance results are provided below;

-   Treatment Group: 24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70

-   Control Group: 42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42

-   Design and execute a two sample t test.\

    a)  State the appropriate null and alternative hypotheses.

    ```{r}
    # H0: There is no relationship between new directed reading activities in the classroom and improvement of some aspects of elementary school pupils' ability.

    # Ha: There is a relationship between new directed reading activities in the classroom and improvement of some aspects of elementary school pupils' ability.
    ```

    b)  Use and show R code to produce the p value and the confidence interval c) Determine if you should reject or fail to reject the null hypothesis using the p value and the confidence interval.

    ```{r}
    treatment <- c(24, 61, 59, 46, 43, 44, 52, 43, 58, 67, 62, 57, 71, 49, 54, 43, 53, 57, 49, 56, 33, 74, 70)
    control <- c(42, 33, 46, 37, 43, 41, 10, 42, 55, 19, 17, 55, 26, 54, 60, 28, 62, 20, 53, 48, 37, 85, 42)
    t_test <- tidy(t.test(treatment, control, mu = 0, var.equal = FALSE))
    print(paste("p-value:",t_test$p.value,"conf.low:",t_test$conf.low,"conf.high:",t_test$conf.high))
    # p-value is less than 0.0105, thus there is enough evidence to reject the null hypothesis. Difference of 0, as suggested by the null hypothesis, also falls outside of the 95% confidence interval of 2.90 to  20.58.
    ```
