---
title: "homework_git_basic_stat"
output: html_document
date: '2023-03-23'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}


```

In Hong Kong, human male height is approximately normally distributed with mean 171.5 cm and standard deviation 5.5 cm. (Use and show R code to produce answers for a – e) You can use any method for part f.

a) What proportion of the Hong Kong population is between 170 cm and 180 cm?

```{r}
l <- 171.5 - 5*5.5
u <- 171.5 + 5*5.5
x <- seq(l, u, length = 100)
y <- dnorm(x = x, mean = 171.5, sd = 5.5)

prob_170 <- pnorm(q = 170, mean = 171.5, sd = 5.5)
prob_180 <- pnorm(q = 180, mean = 171.5, sd = 5.5)
prob_180 - prob_170

```

b) What proportion of the Hong Kong population is less that 165 cm ?

```{r}
prob_165 <- pnorm(q = 165, mean = 171.5, sd = 5.5)
1- prob_165

```


c) What proportion of the Hong Kong population is greater than 175 ?

```{r}
prob_175 <- pnorm(q = 175, mean = 171.5, sd = 5.5)
prob_175

```

d) The proportion .25 is less than what male height value ?


```{r}
x_25 <- qnorm(p = 0.25, mean = 171.5, sd = 5.5)

x_25


```

e) The proportion .85 is greater than what male height value ?

```{r}
x_15 <- qnorm(p = 0.15, mean = 171.5, sd - 5.5)

x_15

```

f) The proportion .68 is between what two male height values?

```{r}
x_16 <- qnorm(p = 0.16, mean = 171.5, sd = 5.5)
x_84 <- qnorm(p = 0.84, mean = 171.5, sd = 5.5)

x_16
x_84

```

2.Use and show R code to determine the height of the standard normal curve at a Z value of 2.5.

```{r}
dnorm(x = 2.5) -> 
   z0
z0

```

3. Use and show R code to determine the height of the standard normal curve at a Z value of .4.
```{r}
dnorm(x = 0.4) -> 
   z0
z0

```

4. Use and show R code to determine the height of a t distribution curve for a t value of 1.5 with 3 degrees of freedom.
```{r}
t_1.5 <- dt(x = 1.5, df = 3)

t_1.5
```

5. Use and show R code to determine the height of a t distribution curve for a t value of -2 with 2 degrees of freedom.
```{r}
dt(-2,2) ->
t_n2

t_n2

```
6. For a t distribution with 3 degrees of freedom, use and show R code that will find the proportion less than 2.5.
```{r}
t_2.5 <- pt(q = 2.5, df = 3)
t_2.5

```

7. For a t distribution with 1 degree of freedom, use and show R code that will find the proportion that is greater than 1.75.

```{r}
1-pt(1.75,1)

```

8. For a t distribution with 2 degrees of freedom, use and show R code that will find the value immediately above a proportion of .355.

```{r}
qt(0.355, df = 2)

```


9. 

a) 

H0: There is no statistical difference between the treatment group and the control group.
Ha: There is a statistical difference between the treatment group and the control group.

b)
```{r}
c(24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70) -> tg
c(42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42) -> cg

t.test(tg,cg, mu = 0, var.equal = FALSE)


```

c)
We can reject the null hypothesis because the p value is lower than the assumed alpha of .05. 
