---
title: "Basic Stat Lab"
author: 'Mark Tenenbaum'
format: html
editor: visual
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3.5, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

# Basic Statistics Classwork/Lab

1)  In Hong Kong, human male height is approximately normally distributed with mean 171.5 cm and standard deviation 5.5 cm. (Use and show R code to produce answers for a -- e) You can use any method for part f.\
    a) What proportion of the Hong Kong population is between 170 cm and 180 cm?

    ```{r}
    mean <- 171.5
    sd <- 5.5

    zupper <- (170 - mean) / sd
    zlower <- (180 - mean) / sd

    pnorm(zlower, mean = 0, sd = 1) - pnorm(zupper, mean = 0, sd = 1)

    ```

    \
    b) What proportion of the Hong Kong population is less that 165 cm ?

    ```{r}
    z <- (165 - mean) / sd
    pnorm(z, mean = 0, sd = 1, lower.tail = TRUE)
    ```

    \
    c) What proportion of the Hong Kong population is greater than 175 ?

    ```{r}
    z <- (175 - mean) / sd
    pnorm(z, mean = 0, sd = 1, lower.tail = FALSE)
    ```

    \
    d) The proportion .25 is less than what male height value ?

    ```{r}
    qnorm(0.25, mean = mean, sd = sd)
    ```

    \
    e) The proportion .85 is greater than what male height value ?

    ```{r}
    qnorm(0.85, mean = mean, sd = sd)
    ```

    \
    f) The proportion of .68 is between what two male height values ?

    ```{r}
    qnorm(0.16, mean = mean, sd = sd)
    qnorm(0.84, mean = mean, sd = sd)
    ```

    \

2)  Use and show R code to determine the height of the standard normal curve at a Z value of 2.5.

```{r}
z <- 2.5
dnorm(z, mean = 0, sd = 1)
```

3)  Use and show R code to determine the height of the standard normal curve at a Z value of .4.

```{r}
z <- 0.4
dnorm(z, mean = 0, sd = 1)
```

4)  Use and show R code to determine the height of a t distribution curve for a t value of 1.5 with 3 degrees of freedom.

```{r}
t <- 1.5
degrees <- 3
dt(t, df = degrees)
```

5)  Use and show R code to determine the height of a t distribution curve for a t value of -2 with 2 degrees of freedom.

```{r}
dt(-2, df = 2)
```

6)  For a t distribution with 3 degrees of freedom, use and show R code that will find the proportion less than 2.5.

```{r}
pt(2.5, df = 3)
```

7)  For a t distribution with 1 degree of freedom, use and show R code that will find the proportion that is greater than 1.75.

```{r}
pt(2.5, df = 3)
```

8)  For a t distribution with 2 degrees of freedom, use and show R code that will find the value immediately above a proportion of .355.

```{r}
qt(0.355, df = 2)
```

9)  An educator believes that new directed reading activities in the classroom will help elementary school pupils improve some aspects of their ability. She arranges for a third -grade class of 23 students to take part in these activities for an eight-week period. A control classroom of 23 third graders follows the same curriculum without the activities. At the end of the eight weeks, all students are given a Degree of Reading Power (DRP) test, which measures the aspects of reading ability that the treatment is designed to improve. The sample data performance results are provided below;

-   Treatment Group: 24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70

-   Control Group: 42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42

-   Design and execute a two sample t test.\
    a) State the appropriate null and alternative hypotheses.

    -   The null hypothesis is that there is no difference in test scores between the treatment groups and the control groups

    -   The alternative hypothesis is that there a difference in test scores between the treatment groups and control groups

        ```{r}
        treatment_group <- c(24, 61, 59, 46, 43, 44, 52, 43, 58, 67, 62, 57, 71, 49, 54, 43, 53, 57, 49, 56, 33, 74, 70)
        control_group <- c(42, 33, 46, 37, 43, 41, 10, 42, 55, 19, 17, 55, 26, 54, 60, 28, 62, 20, 53, 48, 37, 85, 42)

        # Conduct two-sample t-test
        t.test(treatment_group, control_group, alternative = "two.sided")

        ```

        b)  Use and show R code to produce the p value and the confidence interval c) Determine if you should reject or fail to reject the null hypothesis using the p value and the confidence interval

        -   The p-value on the above is 0.0105 (see above output)

        -   The 95% confidence interval is between 2.9 and 20.5
